\section{Acknowledgments}
\vspace{1cm}
\begin{center}
  \begin{minipage}[c]{0.7\linewidth}
    \begin{center}
      \vspace{0.5cm}
      \textit{``Share your knowledge. It is a way to achieve immortality.''} \\
      \hfill --- Dalai Lama
    \end{center}
  \end{minipage}
\end{center}

\vspace{1cm}

I would like to express my sincere gratitude to everyone who supported me during my graduate studies and the completion of this thesis.

I thank the University of Ottawa and the Department of Mathematics and Statistics for providing an excellent environment for my research and studies.

I am deeply grateful to my supervisors, Dr. Maia Fraser and Dr. Vadim Kaimanovich, for their guidance, expertise, and patience throughout this project. Their insights and feedback were invaluable to the development of this work.

My thanks also extend to the faculty members and committee members who offered their knowledge and constructive criticism. I appreciate the time and effort they dedicated to reviewing this thesis.

To my fellow students and colleagues, thank you for the stimulating discussions, collaboration, and friendship that made this journey more enjoyable.

I am grateful to my family and friends for their unwavering support and encouragement during challenging times.

Finally, I acknowledge the financial support from scholarships and teaching assistantships that made this research possible.

Any errors that remain in this work are my own.

\vspace{1cm}
\begin{center}
  \textit{Thank you.}
\end{center}

%%%  Local Variables:
%%%  mode: latex
%%%  TeX-master: "../thesis"
%%%  End: